---
title: "ForMakingPkg"
author: "Zhenyu"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
library(devtools, usethis)
library(Rcpp)
```

```{r}
use_package("RcppParallel")
use_package("RcppXsimd")
use_package("Rcpp")
load_all()
check()
use_mit_license()
usethis::use_rcpp() #you have to run this command in the same directory where src/ is
document()
?rcmg
```

```{r prepare inputs}
#require(MixMatrix)
require(magrittr)
#require(coda)
#require(matrixcalc)
#require(clusterGeneration)

d = 300
mu = rep(0, d)
set.seed(666)
cov_mat = rWishart(1, 2 * d, diag(d))[ , , 1]
cor_mat = cov2cor(cov_mat)
# rcorrmatrix(d, alphad = 1) # from the lkj prior. runs slow though.
# cor_mat = CSgenerate(d, 0.9)
prec = solve(cor_mat)
```

```{r must pass this test - same results as BEAST code}
# compared to the result from BEAST on the SAME target distribution and momentum. BEAST random seed = 666.
# Remember to use the same 1. deterministic 0.1 - 0.9 sequence. 2. step size = 0.1
require(data.table)
require(hzz)
d = 20
n_iter = 1
mu = rep(0, d)

cor_mat = matrix(data = 0.9, nrow = d, ncol = d)
diag(cor_mat) = rep(1, d)
prec = solve(cor_mat)
p0 = rep(0.1, d)

momentum = c(-0.3980153048394741, 2.2483327571774736, -0.2952917086118517, -1.1674703682967922, -0.697182539428704, 1.0133676503829188, -0.21863001118140146, -0.050991178623900166, -0.2829875366316006, 0.6144802843145316, -0.31526635761418426, 2.031161980012812, -0.13072155521386566, 2.306679535948357, -0.5229072378720602, 0.9599761774417773, 0.633641977259469, 2.8769158533009525, -1.755761831952427, -0.600991531699922)

resBeast = c(2.559161,2.644123,2.618147,2.900000,2.787300,2.755389,2.667188,2.715712, 2.627473,2.829323,2.603972,2.696163,2.695163,2.629874,2.595481,2.769309, 2.822751,2.558309,2.900000,2.684678)

res = rcmg(
  n = n_iter,
  mean = mu,
  prec = prec,
  constraits = rep(1, d),
  t = 0.1,
  burnin = 0,
  p0 = p0,
  fixedMomentum = momentum,
  randomFlg = F,
  cppFlg = T,
  nutsFlg = T,
  debug_flg = F
)

# df = as.data.frame(fread(file="/Users/zhenyu/tmp_beast/ZNUTS_20d.log", header = TRUE, quote ="")) 
# df[2, 3:22]
assertthat::assert_that(max(resBeast - res) < 1e-5)
```


```{r to profile}
require(profvis)
#require(hzz)
load_all(export_all=FALSE)
require(tmvtnorm)
set.seed(666)
p0 <- runif(d, 0, 0.2)

t0 = proc.time()

samples_rcpp = rcmg(n = 100, mean = mu, prec = prec, constraits = rep(1, d), t = 10, burnin = 0, p0 = p0, cpp_flg = T,nuts_flg = T, debug_flg = F)
t1 = proc.time()
cat("cpp code takes", t1[3] - t0[3], 'secs')
################################################################
set.seed(666)
p0 <- runif(d, 0, 0.2)
t0 = proc.time()

samples_r = rcmg(n = 100, mean = mu, prec = prec, constraits = rep(1, d), t = 10, burnin = 0, p0 = p0, cpp_flg = F, debug_flg = F)

t1 = proc.time()
cat("r code takes", t1[3] - t0[3], 'secs')
################################################################
```

```{r}
samples_1 = rcmg(n = 30, mean = mu, prec, constraits = rep(1, d), t = 1, burnin = 0, p0 = rep(0.1, d), cpp_flg = FALSE)

profvis({samples_2 = rcmg(n = 3, mean = mu, prec = prec, constraits = rep(1, d), t = 10, burnin = 0, p0 = rep(0.1, d), cpp_flg = T)})

tmvtnorm::rtmvnorm(n = 3, mean = mu, H = prec, lower = rep(0, d), upper = rep(Inf, d), algorithm="gibbs")
```

```{r debug rcpp}
load_all()

get_prec_product <- function (x) {
  if (length(x) == 1){
    return(prec[, x])
  } else {
    return(drop(prec %*% x))
  }
}

set.seed(666)
position <- runif(d, 0, 0.2)
print(position)
set.seed(666)
momentum <-
    (2 * (runif(d) > .5) - 1) * rexp(d, rate = 1)
velocity <- sign(momentum)
gradient <- get_prec_product(position)
action <- get_prec_product(velocity)

engine0 = hzz::createEngine(dimension = d, mask = rep(1, d), observed = rep(1, d), parameterSign = rep(1, d), flags = 1, info = 1, seed = 1)

t0 = proc.time()

t = hzz::.oneIteration(sexp = engine0$engine, position = position, velocity = velocity, action = action, gradient = - gradient, momentum = momentum, time = 20, precision = prec, dimension = d)

t1 = proc.time()
cat("c++ code takes", t1[3] - t0[3], 'secs')

print(t)
```


```{r to compare with ground truth}
require(hzz)
library(MASS)
mu = seq(1,d,1)
n_hzz = 10000
#debugSource("~/zigzag/R/hzz.R", echo=TRUE)
t0 = proc.time()
samples = rcmg(n = n_hzz, mean = mu, prec = prec, constraits = rep(1, d), t = 1, burnin = 0, p0 = p0, randomFlg = T, cppFlg = T, nutsFlg = F, debug_flg = F)

t1 = proc.time()
cat(n_hzz, "hzz samples with dimension", d, "takes", t1[3] - t0[3], 'secs')

n_reject = 10^5
true_samples <- t(mvrnorm(n_reject, mu, cor_mat))
true_samples <- true_samples[, colSums(true_samples < 0) == 0]
print(dim(true_samples))

index <- sample(1:d, 1)

breaks <- seq(0, max(samples, true_samples) + 0.1, length=21)
breaks <- seq(min(samples, true_samples) - 0.1, 0, length=21)
hist(samples[index, ], breaks=breaks, probability=T, col=rgb(1, 0, 0, 1/4), xlim=c(-5, 0))
hist(true_samples[index, ], breaks=breaks, probability=T, col=rgb(0, 0, 1, 1/4), xlim=c(-5, 0), add=T)
```

```{r tmp to fix non-zero mean}
load_all(export_all=FALSE)

d = 10
mu = seq(1,d,1)

prec = diag(d)
cov = matrix(0.9, d, d)
diag(cov) = rep(1, d)
prec = solve(cov)
p0 = rep(0.1, d)

n_hzz = 10000
n_reject = 10^5

samples = rcmg(n = n_hzz, mean = mu, prec = prec, constraits = c(rep(-1, d/2), rep(1, d/2)), t = 10, burnin = 0, p0 = NULL, randomFlg = T, cppFlg = T, nutsFlg = T, debug_flg = F)

true_samples <- t(MASS::mvrnorm(n_reject, mu, cov))
true_samples <- true_samples[, colSums(true_samples < 0) == 0]
dim(true_samples)

samples_tn <- TruncatedNormal::rtmvnorm(n = n_hzz, mu = mu, sigma = cov, lb = c(rep(-Inf, d/2),rep(0, d/2)), ub = c(rep(0, d/2), rep(Inf, d/2)))

plot(samples[1,], samples[2,])
apply(samples, 1, mean)

plot(true_samples[1,], true_samples[2,])
apply(true_samples, 1, mean)


plot(samples_tn[,1], samples_tn[,2])
apply(samples_tn, 2, mean)
```

```{r for debug (not reliable results from old code)}
source("~/truncated-normal-sampler-code/truncated-normal-sampler-code/truncated_normal_hmc.R", echo=TRUE)
f <- function (x) {
  if (length(x) == 1) {
    prec[, x]
  } else {
    drop(prec %*% x)
  }
}

t <- 1
n_samples <- 100
samples <- array(0, c(2, n_samples))

p0 = c(0.1, 0.1)
for (i in 1:n_samples) {
  #debugSource("~/truncated-normal-sampler-code/truncated-normal-sampler-code/truncated_normal_hmc.R", echo=TRUE)
  p0 <- truncated_normal_hmc(f, p0, mu = c(0,0), t = 1, m=NULL)
  samples[, i] <- p0
}
```

I found the following packages that can sample from a multivariate truncated normal:
The results is for our naive version, pure-R code.
### Package tmvtnorm
1. lower dimensions tmv wins clearly. For d = 500, cor = CS(0.9), tmvt takes 10^5 iterations and 100 secs to get min ESS 218. While hzz suffers from "*** recursive gc invocation"(haven't found what this exactly means) with larger stepsize (like 10). And stepsize = 1 takes forever to get to minESS 200. Maybe I should fix the garbage collection in my code?

2. What's the limit (if any) of tmv? 
I tested on CS(0.9) for 
d = 1000: 
(tmv) 885 sec, 2*10^5 iterations, minESS = 370 (tn) 31 secs, 500 iterations, minESS 270
tn wins obviously. And their marginal distributions are quite similar.
d = 2000: CS(0.99) tn takes 500 iterations 210 secs to get 404minESS 
...
Okay I realize (from Botev's paper) tn works great for positive correlated dimensions. 

For a random correlation matrix (eg. from rWishart, it has very low acceptance probability. d = 200 is all right, but d = 300 I cannot get even one single sample.)

Botev mentioned their methods can be a step within a Gibbs sampler to speed it up.

### Package TruncatedNormal
1. CS(0.9) d = 200, piece of cake; d = 500, 1000 iterations, 10 sec, gives min ESS 630, clearly beats tmv.

Note: another pacakge "truncnorm" only works on MVN with independent dimensions.

```{r compare with package truncnorm}
require(hzz)
#require(tmvtnorm)
require(TruncatedNormal)

require(coda)
require(magrittr)

d = 325
mu = seq(1, d, 1)
mu = rep(0, d)

prec = diag(d)
cov = matrix(0.9, d, d)
diag(cov) = rep(1, d)
prec = solve(cov)

n_tn = 1000
n_hzz = 10000

t0 = proc.time()
samples_tn <-
  TruncatedNormal::rtmvnorm(
    n = n_tn,
    mu = mu,
    sigma = cov,
    lb = rep(0, d),
    ub = rep(Inf, d)
  )
t1 = proc.time()

cat(n_tn,
    "tn samples with dimension",
    d,
    "takes",
    t1[3] - t0[3],
    'secs',
    '\n')

mcmc(samples_tn, thin = 1) %>% effectiveSize() %>% summary()

plot(samples_tn[, 1], samples_tn[, 2])


t0 = proc.time()
samples = rcmg(
  n = n_hzz,
  mean = mu,
  prec = prec,
  constraits = rep(1, d),
  t = 1.4,
  burnin = 0,
  p0 = NULL,
  randomFlg = T,
  cppFlg = T,
  nutsFlg = F,
  debug_flg = F
)
t1 = proc.time()
cat(n_hzz,
    "hzz samples with dimension",
    d,
    "takes",
    t1[3] - t0[3],
    'secs',
    '\n')

mcmc(t(samples), thin = 1) %>% effectiveSize() %>% summary()
plot(samples[1,], samples[2,])
```

```{r}

```


```{r continued}
n_tmv = 200000
t0 = proc.time()
samples_tmv <- tmvtnorm::rtmvnorm(n = n_tmv, mean = mu, H = prec, lower = rep(0, d), upper = rep(Inf, d), algorithm="gibbs")
t1 = proc.time()
cat(n_tmv, "tmvn samples with dimension", d, "takes", t1[3] - t0[3], 'secs')

mcmc(samples_tmv, thin = 1) %>% effectiveSize() %>% summary()
# 
# index <- sample(1:d, 1)
# 
# breaks <- seq(0, max(samples_tn, samples_tmv) + 0.1, length=21)
# hist(samples_tn[, index], breaks=breaks, probability=T, col=rgb(1, 0, 0, 1/4), xlim=c(0, 5))
# hist(samples_tmv[, index], breaks=breaks, probability=T, col=rgb(0, 0, 1, 1/4), xlim=c(0, 5), add=T)
```
```{r compare with package TruncatedNormal}

```


