---
title: "for_va_code"
author: "Zhenyu"
date: "2/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#source("~/Dropbox/zigzag/extra/funcsForCompare.R")
require(trialr)
require(rbenchmark)
require(hzz)
require(TruncatedNormal)
require(tmvtnorm)
require(coda)
require(magrittr)
```

# MTN with LKJ-distributed correlations (dimension = 50, 100, 200, 500, )
Use the following chain length to reach a minESS > 200:
(chain)nEXACT, nTN, nHZZ, nNUTS: 800, 200, 1000, 1000


```{r LKJ MTN - prepare inputs}
dimension <- 800 # TN limit is about 300 (endless low acceptance rate)
set.seed(666)
covM <- rlkjcorr(n = 1, K = dimension, eta = 1)
precM <- solve(covM)

meanV <- rep(0, dimension)
constraintForHZZ <- rep(1, dimension)
p0 <- hzz::getInitialValue(mean = meanV, constraits = constraintForHZZ)

lb <- rep(0, dimension)
ub <- rep(Inf, dimension)
```

```{r LKJ MTN - long run to check ESS}
nEXACT <- 10
nTN <- 10
nHZZ <- 10
nNUTS <- 10

res <-  benchMarkTMVN(
  nEXACT = nEXACT,
  nTN = nTN,
  nHZZ = nHZZ,
  nNUTS = nNUTS,
  repeatTimes = 1,
  forOneSampleFlg = F,
  returnSamplesFlg = T,
  dimension = dimension,
  meanVec = meanV,
  precMat = precM,
  covMat = covM,
  pInitial = p0,
  constraintDirec = diag(dimension),
  constraintBdry = rep(0, dimension),
  lbTN = lb,
  ubTN = ub,
  constraintHZZ = constraintForHZZ
)

samplesList <- res$samplesList
cat("dimension is", dimension, "\n")
for (i in 1:length(samplesList)) {
  sampler <- sub(".*samples", "", names(samplesList)[i])
  nName <- paste("n", sampler, sep = '')
  nTotal <- get(nName)
  minESS <- getMinESS(samplesList[[i]])
  n1 <- round(nTotal / minESS, 1)
  cat("minESS (", sampler, ") is", minESS, "n1 is", n1, "---", "n100 is", n1 * 100, "\n")
}
stopifnot("column mean is too different" = max(colMeans(samplesList[[1]]) - colMeans(samplesList[[3]])) < 0.1)
```


n1(d=50)nEXACT, nTN, nHZZ, nNUTS: 1, 1, 2, 3
n100(d=50)nEXACT, nTN, nHZZ, nNUTS: 140, 100, 220, 270

n1(d=100)nEXACT, nTN, nHZZ, nNUTS: 2, 1, 2, 3
n100(d=100)nEXACT, nTN, nHZZ, nNUTS: 170, 100, 240, 250

n1(d=200)nEXACT, nTN, nHZZ, nNUTS: 1, -, 3, 1
n100(d=200)nEXACT, nTN, nHZZ, nNUTS: 140, -, 290, 220

n1(d=400)nEXACT, nTN, nHZZ, nNUTS: 2, -, 4, 3
n100(d=400)nEXACT, nTN, nHZZ, nNUTS: 200, -, 350, 250

n1(d=800)nEXACT, nTN, nHZZ, nNUTS: 2, -, 4, 3
n100(d=800)nEXACT, nTN, nHZZ, nNUTS: 190, -, 350, 300

n1(d=1600)nEXACT, nTN, nHZZ, nNUTS: 2, -, 3, 2
n100(d=1600)nEXACT, nTN, nHZZ, nNUTS: 200, -, 310, 220


```{r LKJ MTN - timing run to get result}
nEXACT <- 190
nTN <- 0
nHZZ <- 350
nNUTS <- 300

res <-  benchMarkTMVN(
  nEXACT = nEXACT,
  nTN = nTN,
  nHZZ = nHZZ,
  nNUTS = nNUTS,
  repeatTimes = 2,
  forOneSampleFlg = T,
  returnSamplesFlg = F,
  dimension = dimension,
  meanVec = meanV,
  precMat = precM,
  covMat = covM,
  pInitial = p0,
  constraintDirec = diag(dimension),
  constraintBdry = rep(0, dimension),
  lbTN = lb,
  ubTN = ub,
  constraintHZZ = constraintForHZZ
)

dplyr::bind_rows(res$timeList)
```

```{r time for 1}
nEXACT <- 2
nTN <- 2
nHZZ <- 2

res <- benchMarkTMVN(nEXACT = nEXACT, nHZZ = nHZZ, repeatTimes = 3)
minESSEXACT <- getMinESS(res$samplesEXACT)
minESSHZZ <- getMinESS(res$samplesHZZ)
minESSNUTS <- getMinESS(res$samplesNUTS)

# timeRes$nFor1Sample <- nTotal / minESS
# timeRes$nFor100Samples <- timeRes$nFor1Sample * 100
# 
# 
# timeRes[timeRes$test == "lkjEXACT", "relative"] <-
#   minESSEXACT / timeRes[timeRes$test == "lkjEXACT", "relative"]
# timeRes[timeRes$test == "lkjTN", "relative"] <-
#   nTN / timeRes[timeRes$test == "lkjTN", "relative"]
# timeRes[timeRes$test == "lkjHZZ", "relative"] <-
#   minESSHZZ / timeRes[timeRes$test == "lkjHZZ", "relative"]
# timeRes[timeRes$test == "lkjNUTS", "relative"] <-
#   minESSNUTS / timeRes[timeRes$test == "lkjNUTS", "relative"]

#print(timeRes$relative)
cat("dimension is", dimension, "\n",
  "min ESS EXACT, HZZ, NUTS:", c(minESSEXACT, minESSHZZ, minESSNUTS), "\n")
print(res$timeRes)


#stopifnot("column mean is too different" = max(colMeans(samplesEXACT) - colMeans(samplesHZZ)) < 0.2)
```


```{r tmp check invgamma distribution}
hist(rinvgamma(n = 1000, shape = 1, rate = 1), xlim = c(0,20), breaks=1000)
#use https://www.r-bloggers.com/2017/05/5-ways-to-measure-running-time-of-r-code/
```

```{r tmp lancoz}
require(RSpectra)
require(trialr)

set.seed(666)
covMat <- rlkjcorr(n = 1, K = 10, eta = 1)
tmp = RSpectra::eigs(A = covMat, k = 1, which = "LM")[["values"]]
tmp$values

```

```{r phylogenetics example}
# read in data
data_folder <- "/Users/zhenyuzhang/probitHIVdata/"
orthant_ind <- scan(paste0(data_folder, "orthant_indicator.txt"))
mean <- scan(paste0(data_folder, "mean.txt"))

tic <- proc.time()

# PrecTotal <- read.table(
#   paste0(data_folder, "precision_matrix.csv"), 
#   header = FALSE, sep = ",", colClasses = "numeric"
# )
toc <- proc.time()
print(sprintf("Loading the prec matrix took %.3g sec.", toc['elapsed'] - tic['elapsed']))

Prec <- as.matrix(PrecTotal)
subsample_data <- TRUE
set.seed(666)

if (subsample_data) {
  subset_ind <- sample.int(dim(Prec)[1], size = 5000)
  orthant_ind <- orthant_ind[subset_ind]
  mean <- mean[subset_ind]
  Prec <- Prec[subset_ind, subset_ind]
}

constraint_directions <- diag(orthant_ind)
constraint_directions <- constraint_directions[!is.na(orthant_ind), ]
constraint_bdry <- rep(0, dim(constraint_directions)[1])

x0 <- runif(length(orthant_ind))
is_not_na <- !is.na(orthant_ind)
x0[is_not_na] <- orthant_ind[is_not_na] * abs(x0[is_not_na])


nEXACT = 50

samplesEXACT <- exactHMC(
          n = nEXACT,
          initial_position = x0,
          constraint_direc = constraint_directions,
          constraint_bound = constraint_bdry,
          precisionMat = Prec,
          mu = mean,
          precision = TRUE,
          total_time = pi / 2,
          seed = 666
        )
```



```{r benchmarking}
require(hzz)

# require(pryr)
# require(devtools)
# load_all()

n <-  10 # number of samples for the probit model
num_samples_TN <-  300 # number of TN samples (independent)
num_samples_hzz  <-  400 # number of hzz samples
# read in data (2011 rows, 10 continuous covariates and 1 binary outcome)
df_water <-
  read.csv(file = "../data/water_processed.csv", header = T)

set.seed(666)
df_water  <-  df_water[sample(1:nrow(df_water), n), ]

X  <-  as.matrix(df_water[,-ncol(df_water)])
X  <-  scale(X)
y_raw  <-  df_water$Potability
y  <-  y_raw
y[y == 0]  <-  -1

p  <-  ncol(X)
n  <-  nrow(X)
dim  <-  p + n

pc  <-  hzz::getPrecCov(X, n, p, sigma2 = 2)

TN_lb  <-  y
TN_ub  <-  y
TN_lb[TN_lb == -1]  <-  -Inf
TN_lb[TN_lb == 1]  <-  0
TN_ub[TN_ub == -1]  <-  -0
TN_ub[TN_ub == 1]  <-  Inf
```

```{r timing}
require(hzz)
require(TruncatedNormal)
require(invgamma)
require(rbenchmark)

samples_hzz  <-  array(0, c(num_samples_hzz, dim))
samples_TN <- array(0, c(num_samples_TN, dim))

meanV <- rep(0, dim)

mask <- rep(1, dim)
observed <- rep(1, dim)
parameterSign <- c(rep(0, p), y)

sample_hzz <-
  hzz::getInitialValue(mean = rep(0, dim), constraits = parameterSign)
sample_TN <- sample_hzz

timeRes <- benchmark(
  "TN" = {
    set.seed(666)
    for (i in 1:num_samples_TN) {
      sigma2 <-
        rinvgamma(
          n = 1,
          shape = 2 + dim / 2,
          scale = 2 + sample_TN %*% sample_TN / 2
        )
      pc <- getPrecCov(X, n, p, sigma2 = sigma2)
      
      sample_TN <- TruncatedNormal::rtmvnorm(
        n = 1,
        mu = meanV,
        sigma = pc$covMat,
        lb = c(rep(-Inf, p), TN_lb),
        ub = c(rep(Inf, p), TN_ub)
      )
      samples_TN[i,] <- sample_TN
    }
  },
  "hzz" = {
    engine = createEngine(
      dimension = dim,
      mask = mask,
      observed = observed,
      parameterSign = parameterSign,
      flags = 128,
      info = 1,
      seed = 666
    )
    
    set.seed(666)
    for (i in 1:num_samples_hzz) {
      sigma2 <-
        rinvgamma(
          n = 1,
          shape = 2 + dim / 2,
          scale = 2 + sample_hzz %*% sample_hzz / 2
        )
      pc <- getPrecCov(X, n, p, sigma2 = sigma2)
      
      setMean(sexp = engine$engine, mean = rep(0, dim))
      setPrecision(sexp = engine$engine, precision = pc$precMat)
      
      momentum <- drawMomentum(dim)
      sample_hzz <- getSample(
        position = sample_hzz,
        momentum = momentum,
        t = 10,
        nutsFlg = F,
        engine = engine
      )
      
      samples_hzz[i, ] <- sample_hzz
    }
  },
  replications = 1,
  columns = c(
    "test",
    "replications",
    "elapsed",
    "relative",
    "user.self",
    "sys.self"
  )
)

print(timeRes)
```


```{r check ESS}
require(coda)
require(magrittr)
mcmc(samples_hzz, thin = 1) %>% effectiveSize() %>% summary()
mcmc(samples_TN, thin = 1) %>% effectiveSize() %>% summary()
colMeans(samples_hzz[,1:9])
colMeans(samples_TN[,1:9])

summary(colMeans(samples_hzz[,1:9]))
summary(colMeans(samples_hzz) - colMeans(samples_TN))
```


```{r}
hist(samples_hzz[,29], breaks = 100)
hist(samples_TN[,29], breaks = 100)
```


```{r UPG}
df_water = read.csv(file = "../data/water_potability.csv", header = T)
colnames(df_water)
df_water = df_water[complete.cases(df_water), ]
myprobit <- glm(Potability ~ Sulfate + Solids + Organic_carbon, family = binomial(link = "probit"), data = df_water)
summary(myprobit)

require(UPG)
data("lfp", package = "UPG")
head(lfp, 5)
y <- c(rep(1, 100), rep(0, 1911))
X <- matrix(rnorm(nrow(df_water) * 200, 0 ,1), nrow = nrow(df_water), ncol = 200)
results.logit <- UPG(y = y, X = X, type = "probit", BOOST = F)
UPG.Diag(results.logit)
summary(results.logit)
```

```{r test new boundary interface}
require(hzz)
d = 4
t = rcmg(
  n = 10,
  mean = rep(0, d),
  cov = NULL,
  prec = diag(d),
  constraits = rep(1, d),
  lowerBounds = rep(0, d),
  upperBounds = rep(Inf, d),
  t = 1,
  # burnin=1,
  p0 = rep(0.1, d),
  fixedMomentum = NULL,
  cppFlg = TRUE,
  nutsFlg = T,
  random_seed = 666,
  randomFlg = T,
  debug_flg = F
)
t_new = t
```

